High-Performance Language Primitives for Distributed Architectures using Asynchronous Events

### 2012-07-23 doc/asplos2013/micro.tex
small edits to the microbenchmarks

Figure~\ref{fig:eventthroo} shows the event trigger rates achieved by our implementation
for a variety of node counts and fan-in/out factors.
For small fan-in/out factors, the total rate falls off initially going to two nodes as active
message become necessary, but starts increasing slightly again at larger node counts.  Higher
fan-in/out factors require more messages and have lower throughput that also increases with
node count.  Although the number of events waiting on each node decreases with an increasing
node code, the minimal scaling indicates that the main bottleneck is in the processing of the
active message that each node must receive rather than the local redistribution of the triggering
notification.
%Note we can further improve performance by increasing
%the number of threads per node for handling active messages (in these experiments there
%is one per node), but in real applications these threads would consume hardware cores and
%take away computational resources from the primary application.  This is only a viable
%option for memory or communication bound programs.

@@

To more clearly demonstrate the benefit of lock unfairness, even at large node counts, we show an alternative
cut of the experiment space in Figure~\ref{fig:fixednode}.
In this case the node count is fixed at 8 and lock grant rates are shown for a variety of total lock counts
and number of chains per node.
With only 32 chains per node (a total of 256 chains)
there is little contention and the grant rate is high.  As the number of chains per node increases there is
more contention for locks and the rate drops as a result.  For smaller lock counts, further increases in
the number of chains result in improved rates.  Note that the point on each line where this occurs
 is when
the chains per node is greater than total locks, i.e. where the expected number of requests per lock 
per node becomes more than one.  As soon as there are multiple expected requestors for the same lock on
a node, the unfairness property of our implementation is able to reduce the number of lock ownership 
transfers, yielding better performance.

@@

\subsection{Reduction Throughput}
\label{subsec:reducmicro}
To show the benefits of associating reduction operations with physical regions we designed
a histogram micro-benchmark that requires all nodes to perform an addition reduction to a
physical region located in the globally visible GASNet memory (described in 
Section~\ref{sec:impl}).  Using the reduction interface described in Section~\ref{subsec:phyreg}
reductions are performed in five ways:

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Single Reads/Writes - Each node takes a lock on the global physical region and then for each reduction
reads a value from the physical region, performs the addition, and then writes the result back.  The lock is then released.
\item Single Reductions - Each reduction is individually sent as an active message to the node whose
system memory holds the target of the reduction.
\item Localized Instance - Each node takes a lock on the the global physical region, copies it to a new physical
region in its own system memory, performs all reductions to the localized region, copies the region back, and releases the lock.
\item Fold Instance - Each node creates a local reduction instance that folds reductions to the same element.
All reductions are folded into the instance, which is then copied back to the global region.
\item List Instance - Each node creates a local reduction instance that stores a list of reduction operations.
All reductions are appended to the list, which is then copied back to the global region.
\end{itemize}

We ran two experiments, one corresponding with dense reductions and the other corresponding with sparse
reductions.  In each case a large random source of data is divided across into chunks, which are given to
separate reduction tasks.  There are 8 reduction tasks created for each node, one per processor.  The dense
case uses a histogram with 256K buckets and each reduction task processes a chunk of 4M entries (i.e.
performs 4M reductions).  The results for the
dense experiment are shown in Figure~\ref{fig:reducdense}.  The sparse case use a histogram with 4M buckets,
but only 64K elements are processed by each task.  Results for the sparse case
be seen in Figure~\ref{fig:reducsparse}.

The dense experiment demonstrates that reduction fold regions perform 
best and scale well with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  List regions also scale well, but perform about an
order of mangitude worse than folding regions in the dense case.  The use of a separate active message for
each reduction operation is another two orders of magnitude worse, demonstrating that data must be
transferred in larger blocks to be efficient on modern interconnect hardware.  The localized instance approach
works well for a single node, but its inherently serial nature doesn't benefit at all from increasing node counts.  Finally, the only time the latency of performing individual RMA reads and writes isn't disastrous is
on a single node (where the reads and writes aren't really remote).

>>>>

Figure~\ref{fig:eventthroo} shows the event trigger rates achieved by our implementation
for a variety of node counts and fan-in/out factors.
For small fan-in/out factors, the total rate falls off initially going to two nodes as active
messages become necessary, but starts increasing slightly again at larger node counts.  Higher
fan-in/out factors require more messages and have lower throughput that also increases with
node count.  Although the number of events waiting on each node decreases with an increasing
node code, the minimal scaling indicates that the main bottleneck is in the processing of the
active message that each node must receive rather than the local redistribution of the triggering
notification.
%Note we can further improve performance by increasing
%the number of threads per node for handling active messages (in these experiments there
%is one per node), but in real applications these threads would consume hardware cores and
%take away computational resources from the primary application.  This is only a viable
%option for memory or communication bound programs.

@@

To more clearly demonstrate the benefit of lock unfairness, even at large node counts, we show an alternative
cut of the experiment space in Figure~\ref{fig:fixednode}.
In this case the node count is fixed at 8 and lock grant rates are shown for a variety of total lock counts
and number of chains per node.
With only 32 chains per node (a total of 256 chains)
there is little contention and the grant rate is high.  As the number of chains per node increases there is
more contention for locks and the rate drops as a result.  For smaller lock counts, further increases in
the number of chains result in improved rates.  Note that the point on each line where this occurs
is when the chains per node is greater than total locks, which corresponds to the point where the expected 
number of requests per lock 
per node becomes more than one.  As soon as there are multiple expected requestors for the same lock on
a node, the unfairness property of our lock implementation is able to reduce the number of lock ownership 
transfers, yielding better performance.

@@

\subsection{Reduction Throughput}
\label{subsec:reducmicro}
To show the benefits of associating reduction instances we designed
a histogram micro-benchmark that requires all nodes to perform an addition reduction to a
physical region located in the globally visible GASNet memory (described in 
Section~\ref{sec:impl}).  Using the reduction interface described in Section~\ref{subsec:phyreg}
reductions are performed in five ways:

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Single Reads/Writes - Each node takes a lock on the global physical region and then for each reduction
reads a value from the physical region, performs the addition, and then writes the result back.  The lock is then released.
\item Single Reductions - Each reduction is individually sent as an active message to the node whose
system memory holds the target of the reduction.
\item Localized Instance - Each node takes a lock on the the global physical region, copies it to a new physical
region in its own system memory, performs all reductions to the localized region, copies the region back, and releases the lock.
\item Fold Instance - Each node creates a local fold reduction instance.
All reductions are folded into the instance, which is then copied back to the global region.
\item List Instance - Each node creates a local list reduction instance.
All reductions are appended to the list reduction instance, which is then copied back to the global region.
\end{itemize}

We ran two experiments, one corresponding with dense reductions and the other corresponding with sparse
reductions.  In each case a large random source of data is divided across into chunks, which are given to
separate reduction tasks.  There are 8 reduction tasks created for each node, one per processor.  The dense
case uses a histogram with 256K buckets and each reduction task performs 4M reductions.  The results for the
dense experiment are shown in Figure~\ref{fig:reducdense}.  The sparse case use a histogram with 4M buckets,
but only 64K reductions are performed by each task.  Results for the sparse case
be seen in Figure~\ref{fig:reducsparse}.

The dense experiment demonstrates that reduction fold regions perform 
best and scale well with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  List regions also scale well, but perform about an
order of mangitude worse than folding regions in the dense case.  The use of a separate active message for
each reduction operation is another two orders of magnitude worse, demonstrating that data must be
transferred in larger blocks to be efficient on modern interconnect hardware.  The localized instance approach
works well for a single node, but its inherently serial nature doesn't benefit at all from increasing node counts.  Finally, the only time the latency of performing individual RMA reads and writes isn't disastrous is
on a single node (where the reads and writes are local).
